Architectures and Open pNets

### 2018-10-18 2018-Arch-pNets/arch-pNets.tex
Arch-pNet paper: some modifications in response to comments by Ludo: Expanded the explanation of how the tick-fail-fail connector in the running example defines three interactions

Intuitively, we 1)~introduce a fresh initial state with an additional
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$ and 2)~define a pNet action algebra that associates, as action parameter, an
additional Boolean value to each port, 3)~create pNet transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 4)~create additional loop transitions marked by
$\false$.

@@

The encoding of the interaction model is based on its representation
as a set of connectors. \noteSB{A legitimate question is whether we
  actually need to do this, instead of listing explicitly the
  individual interactions, then associating a synchronisation vector
  to each one of them?  One of the factors is the encoding
  pNets$\rightarrow$SAT: is it better to have more or fewer SVs if the
  price is additional Boolean variables?  The other factor might be
  the facility of maximal progress encoding\mdash TBC.  Traceability.
  Anything else?  }Indeed, as illustrated by the Failure Monitor
architecture in \fig{schema:ArchFailure:BIP}, each connector can
define several allowed interactions, depending on its hierarchical
structure and port and node syncrhon/trigger typings.  For instance,
the connector {\NameTimer.\PortTick \trigsynch (\PortFail \trigsynch
  \NameCtrl.\PortFail)} defines three interactions (observe that
$\topinterval + \NameTimer.t = \topinterval$ and $\NameTimer.z \cap
\topinterval = \NameTimer.z$):\noteSB{Move this to \secn{components}?}
%
\begin{align*}
  &\bigl(\{\NameTimer.\PortTick\},
  &&\true, 
  &&\NameTimer.z := \NameTimer.z\bigr)
  \\
  &\bigl(\{\NameTimer.\PortTick, \PortFail\},
  &&\true,
  &&\NameTimer.z := \NameTimer.z\bigr)
  \\
  &\bigl(\{\NameTimer.\PortTick, \PortFail, \NameCtrl.\PortFail\},
  &&\true,
  &&\NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
\end{align*}

In order to encode these using only one synchronisation vector, we use
the additional Boolean values associated to each port by the encoding
of coordinator components.  We observe that the three ports in this
connector form a ``causality chain'': {\NameCtrl.\PortFail} can only
participate in the interaction if the dangling port {\PortFail}
participates, which in turn can only participate if
{\NameTimer.\PortTick} does so.  These dependencies can be rewritten
as Boolean implications $\NameCtrl.\PortFail \Rightarrow \PortFail$
and $\PortFail \Rightarrow \NameTimer.\PortTick$.  However, notice
that, within the scope of this connector, {\NameTimer.\PortTick}
participates in all the interactions, meaning that the latter
implication is a tautology and can be discarded.  Thus, we obtain the
synchronisation vector SV0 shown in \fig{enc:nopri}, where $b_0$ and
$b_1$ are the Boolean values associated to the actions encoding the
ports {\PortFail} and {\NameCtrl.\PortFail}, $\alpha_0$ is a fresh
action name.  The guard $b_1 \Rightarrow b_0$ encodes the causal
relation among ports {\NameCtrl.\PortFail} and {\PortFail}.  Notice
that, in our encoding, all three ports are present in the
synchronisation vector.  Non-participation of a port corresponds to
firing the loop transition marked by \false.  \Fig{enc:nopri} shows
the four synchronisation vectors SV0\ndash SV3 corresponding to the
connectors in \fig{schema:ArchFailure:BIP} and an additional vector
SV4, synchronising the {\init} transitions of the two pLTSs.

>>>>

Intuitively, we 1)~introduce a fresh initial state with an additional
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~define a pNet action algebra that associates, as action parameter, an
additional Boolean value to each port, 3)~create pNet transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 4)~create additional loop transitions marked by
$\false$.

@@

The encoding of the interaction model is based on its representation
as a set of connectors. \noteSB{%
%
  A legitimate question is whether we actually need to do this,
  instead of listing explicitly the individual interactions, then
  associating a synchronisation vector to each one of them?  One of
  the factors is the encoding pNets$\rightarrow$SAT: is it better to
  have more or fewer SVs if the price is additional Boolean variables?
  The other factor might be the facility of maximal progress
  encoding\mdash TBC.  Traceability.  Anything else?
%
} Indeed, as illustrated by the Failure Monitor architecture in
\fig{schema:ArchFailure:BIP}, each connector can define several
allowed interactions, depending on its hierarchical structure and the
use of syncrhons and trigger.  \noteSB{Move this to
  \secn{components}?}For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} is a
two-level hierarchical connector.  In the second-level subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  Therefore, this
subconnector allows two interactions: one with {\PortFail} alone and
another one involving both ports.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, whereas the second-level
subconnector is a synchron.  Hence, this connector allows an
interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%
  Such interactions are called \emph{singleton interactions}.
%
} and any interaction involving both {\NameTimer.\PortTick} and
\emph{some} interaction of the second-level subconnector.  Combining
the two levels we can see that the entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
\begin{align*}
  &\bigl(\{\NameTimer.\PortTick\},
  &&\true, 
  &&\NameTimer.z := \NameTimer.z\bigr)
  \\
  &\bigl(\{\NameTimer.\PortTick, \PortFail\},
  &&\true,
  &&\NameTimer.z := \NameTimer.z\bigr)
  \\
  &\bigl(\{\NameTimer.\PortTick, \PortFail, \NameCtrl.\PortFail\},
  &&\true,
  &&\NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
\end{align*}

In order to encode these interactions using only one synchronisation vector, we use
the additional Boolean values associated to each port by the encoding
of coordinator components.  We observe that the three ports in this
connector form a ``causality chain'': {\NameCtrl.\PortFail} can only
participate in the interaction if the dangling port {\PortFail}
participates, which in turn can only participate if
{\NameTimer.\PortTick} does so.  These dependencies can be rewritten
as Boolean implications $\NameCtrl.\PortFail \Rightarrow \PortFail$
and $\PortFail \Rightarrow \NameTimer.\PortTick$.  However, notice
that, within the scope of this connector, {\NameTimer.\PortTick}
participates in all the interactions, meaning that the latter
implication is a tautology and can be discarded.  Thus, we obtain the
synchronisation vector SV0 shown in \fig{enc:nopri}, where $b_0$ and
$b_1$ are the Boolean values associated to the actions encoding the
ports {\PortFail} and {\NameCtrl.\PortFail}, $\alpha_0$ is a fresh
action name.  The guard $b_1 \Rightarrow b_0$ encodes the causal
relation among ports {\NameCtrl.\PortFail} and {\PortFail}.  Notice
that, in our encoding, all three ports are present in the
synchronisation vector.  Non-participation of a port corresponds to
firing the loop transition marked by \false.  \Fig{enc:nopri} shows
the four synchronisation vectors SV0\ndash SV3 corresponding to the
connectors in \fig{schema:ArchFailure:BIP} and an additional vector
SV4, synchronising the {\init} transitions of the two pLTSs.
